---
title: 'STOR 455 Homework 7'
subtitle: '20 points - Due Wednesday 4/3 before class'
output:
  pdf_document: default
---

### Are Emily and Greg More Employable Than Lakisha and Jamal? ###

Bertrand, M., & Mullainathan, S. (2004). Are Emily and Greg more employable than Lakisha and Jamal? A field experiment on labor market discrimination. _American Economic Review, 94_(4), pp. 991-1013.

\[Abstract\]
We perform a field experiment to measure racial discrimination in the labor market. We respond with fictitious resumes to help-wanted ads in Boston and Chicago newspapers. To manipulate perception of race, each resume is randomly assigned either a very African American sounding name or a very White sounding name. The results show significant discrimination against African-American names: White names receive 50 percent more callbacks for interviews. We also find that race affects the benefits of a better resume. For White names, a higher quality resume elicits 30 percent more callbacks whereas for African Americans, it elicits a far smaller increase. Applicants living in better neighborhoods receive more callbacks but, interestingly, this effect does not differ by race. The amount of discrimination is uniform across occupations and industries. Federal contractors and employers who list “Equal Opportunity Employer” in their ad discriminate as much as other employers. We find little evidence that our results are driven by employers inferring something other than race, such as social class, from the names. These results suggest that racial discrimination is still a prominent feature of the labor market.    


Variables     | Descriptions
-----------   | -----------------------------------------------------
_call_        | Was the applicant called back? (1 = yes; 0 = no)
_ethnicity_   | indicating ethnicity (i.e., "Caucasian-sounding" vs. "African-American sounding" first name)
_sex_         | indicating sex
_quality_     | Indicating quality of resume.
_experience_  | Number of years of work experience on the resume
_equal_       | Is the employer EOE (equal opportunity employment)?


Use the _ResumeNames455_ found at the address below:

https://raw.githubusercontent.com/JA-McLean/STOR455/master/data/ResumeNames455.csv

```{r}
library(tidyverse)
library(Stat2Data)

EmployData <- read_csv("https://raw.githubusercontent.com/JA-McLean/STOR455/master/data/ResumeNames455.csv", show_col_types = FALSE)
```

1) An Equal Opportunity Employer (EOE) is an employer who agrees not to discriminate against any employee or job applicant because of race, color, religion, national origin, sex, physical or mental disability, or age. Construct a logistic model to predict if the job applicant was called back using _ethnicity_, _equal_, _sex_, and the interactions between _ethnicity_ and _equal_, and _sex_ and _equal_ as the predictor variables. 

```{r}
logisticmod1 <- glm(call ~ ethnicity + equal + sex + ethnicity*equal + sex*equal, family = binomial, data = EmployData)
summary(logisticmod1)
```


2)  Conduct a drop in deviance hypothesis test to determine the effectiveness of the _equal_ terms in the model constructed in the previous question. Cite your hypotheses, p-value, and conclusion in context.

```{r}
reducedmod <- glm(call ~ ethnicity + sex, family = binomial, data = EmployData)
1 - pchisq(summary(reducedmod)$deviance - summary(logisticmod1)$deviance, 3)
```

H0: The _equal_ terms have a coefficient of 0 (they do not significantly improve the model). H1: The _equal_ terms have a non-zero coefficient (they significantly improve the model). Our p-value here is 0.2628813, which is larger than 0.05. Thus, we fail to reject the null. There is not strong enough evidence that the _equal_ terms significantly improve the model. Thus, there is not a significant difference between the full model (logisticmod1) and the reduced model.

3)  Based on your model from question 1, What is the probability of a female applicant with a "Caucasian-sounding" name getting a call back from an Equal Opportunity Employer (EOE)? What is the probability of a male applicant with an "African-American sounding" name getting a call back from an Equal Opportunity Employer (EOE)?

```{r}
cauc_woman <- data.frame(sex = "female", ethnicity = "cauc", equal = "yes")
predict(logisticmod1, cauc_woman, type = 'response')
```

According to this model, the probability of a female applicant with a "Caucasian-sounding" name getting a call back from an Equal Opportunity Employer is around 9.81%.

```{r}
afam_man <- data.frame(sex = "male", ethnicity = "afam", equal = "yes")
predict(logisticmod1, afam_man, type = 'response')
```

According to this model, the probability of a male applicant with an "African-American sounding" name getting a call back from an Equal Opportunity Employer is around 6.96%.

4) Does the number of years of work experience impact the relationship between _ethnicity_, _sex_, and an applicant getting called back? Construct a logistic model to predict if the job applicant was called back using _ethnicity_, _sex_, _experience_, and the interactions between _ethnicity_ and _experience_, and _sex_ and _experience_ as the predictor variables. 

```{r}
logisticmod2 <- glm(call ~ ethnicity + experience + sex + ethnicity*experience + sex*experience, family = binomial, data = EmployData)
summary(logisticmod2)
```

5)  Conduct a drop in deviance hypothesis test to determine the effectiveness of the _experience_ terms in the model constructed in the previous question. Cite your hypotheses, p-value, and conclusion in context.

```{r}
reducedmod2 <- glm(call ~ ethnicity + sex, family = binomial, data = EmployData)
1 - pchisq(summary(reducedmod2)$deviance - summary(logisticmod2)$deviance, 3)
```

H0: The _experience_ terms have a coefficient of 0 (they do not significantly improve the model). H1: The _experience_ terms have a non-zero coefficient (they significantly improve the model). Our p-value here is 4.957569e-05, which is smaller than 0.05. Thus, we reject the null. There is strong evidence that the _experience_ terms significantly improve the model. Thus, there is a significant difference between the full model and the reduced model. Thus, we can likely conclude that the number of years of work experience do impact the relationship between _ethnicity_, _sex_, and an applicant getting called back, as it seems that adding the _experience_ terms significantly improves the model.

6)  Use an appropriate model selection method to construct a best model to predict if the job applicant was called back using any of the variables as predictors (except for _name_). You do not need to consider interaction terms. Why would you not want to use _name_ as a predictor?

```{r}
EmployData1 <- select(EmployData, -c(name))
full_logisticmod1 <- glm(call ~ ., family = binomial, data = EmployData1)
MASS::stepAIC(full_logisticmod1, trace = FALSE)
```
```{r}
finalmod <- glm(formula = call ~ ethnicity + quality + experience, 
                family = binomial, data = EmployData1)
summary(finalmod)
```

I created my final model using backwards regression, as you can see above.

We would not want to use _name_ as a predictor because there are many different names - in theory, if we were to treat it as a categorical variable, there would be as many "categories" as there are observations in the data, since every resume has a different name. Thus, this would not be helpful in a model at all, as we are not concerned with how a specific name might affect an applicant's chances of being called back. For example, we don't want to know if John has a greater chance of being called back than Bryan. Rather, we are concerned with whether it the name sounds African-American or white, and this is already captured by _ethnicity_.
